Ultrafast Electron Microscopes: Design Criteria, Electron Sources, and Column Modeling

### 2013-04-15 inc/ms_model/ms_model.tex
more tweaks including adding a laser section to the conclusions
typos (Jeff Berger++)

This chapter introduces the Analytic Gaussian (AG) model of Michalik and Sipe~\cite{michalik_analytic_2006}.
In the AG model the dynamics of a Gaussian electron pulse are computed using a mean-field self-similar approximation.
The result is a set of equations that can be solved very quickly using numerical methods on a modern computer.
When checked against N-body simulations, the authors find that the model is also surprisingly accurate, even for a large set of non-Gaussian pulses~\cite{michalik_analytic_2006,michalik_evolution_2009}.

A drawback of this model is that the only force included is that of the internal repulsive space-charge field.
In Chapter \ref{chap:extension}, I will extend this model, adding the influence of external forces, making it more generally applicable for UEM.

@@

For a classical set of $N$ discreet particles (labeled by the index $i$), distributed in 6D space (3 position, 3 momentum) the distribution function is 
\begin{equation}
  f_{\smallD}(\vec{r}, \vec{p}; t) = 
  \sum\limits^{N}_{i=1} \delta(\vec{r}_{i} - \vec{r}) \delta(\vec{p}_{i} - \vec{p}) \,\text{,}
\end{equation}
which automatically satisfies \ref{eq:normalization}.
If the particles have mass $m$ which interact with some potential $V(\vec{r}_{i} - \vec{r}_{j})$, their motion is governed by the equations
\begin{subequations}\label{eq:std_motion}
\begin{gather}
  \frac{d\vec{r}_{i}(t)}{dt} = \frac{\vec{p}_{i}(t)}{m}\\
  \frac{d\vec{p}_{i}(t)}{dt} = -\sum\limits_{j} \frac{d}{d\vec{r}_{i}} V(\vec{r}_{i} - \vec{r}_{j}) \label{eq:std_motion_v} \,\text{.}
\end{gather}
\end{subequations}
Now let us consider the time evolution of $f_{\smallD}$;
\begin{equation}
  \frac{\partial f_{\smallD}}{\partial t} = 
  \frac{d\vec{r}}{dt} \frac{\partial f_{\smallD}}{\partial \vec{r}} 
  + \frac{\partial f_{\smallD}}{\partial \vec{p}} \frac{d\vec{p}}{dt} \,\text{,}
\end{equation}
but with \ref{eq:std_motion} the final discreet evolution equation is
\begin{equation} \label{eq:evolution_discreet}
  \frac{d}{dt} f_{\smallD}(\vec{r}, \vec{p}; t) =
  -\frac{\vec{p}}{m} \frac{\partial}{\partial \vec{r}} f_{\smallD}(\vec{r}, \vec{p}; t)
  + \frac{\partial}{\partial \vec{p}} f_{\smallD} (\vec{r}, \vec{p}; t)
  \frac{\partial}{\partial \vec{r}} \iint d\vecprime{r} d\vecprime{p} V(\vec{r} - \vecprime{r}) f_{\smallD}(\vecprime{r},\vecprime{p};t) \,\text{,}
\end{equation}
where the extra negation comes from the change of perspective of \ref{eq:std_motion_v} from the frame of the individual particle $i$ to the observer frame.

This analysis can be extended to continuous distributions by taking the ensemble average over \ref{eq:evolution_discreet}.
In order to simplify the resulting equation to a useful result, one applies the usual mean-field approximation
\begin{equation}
  \left < f_{\smallD} (\vec{r}, \vec{p}; t) f_{\smallD} (\vecprime{r}, \vecprime{p}; t) \right > = \left < f_{\smallD} (\vec{r}, \vec{p}; t) \right > \left < f_{\smallD} (\vecprime{r}, \vecprime{p}; t) \right > \,\text{,}
\end{equation}
so that the formal evolution of a continuous distribution is given by
\begin{equation} \label{eq:formal_evolution}
  \frac{d}{dt} f(\vec{r}, \vec{p}; t) =
  -\frac{\vec{p}}{m} \frac{\partial}{\partial \vec{r}} f(\vec{r}, \vec{p}; t)
  + \frac{\partial}{\partial \vec{p}} f (\vec{r}, \vec{p}; t)
  \frac{\partial}{\partial \vec{r}} \iint d\vecprime{r} d\vecprime{p} V(\vec{r} - \vecprime{r}) f(\vecprime{r},\vecprime{p};t) \,\text{.}
\end{equation}
Although this equation looks very similar to \ref{eq:evolution_discreet}, notice that it contains approximations, mean-field that has been seen and in practice the self-similarity of the distribution function, that the discreet description did not.

>>>>

This chapter introduces the Analytic Gaussian (AG) model of Michalik and Sipe~\cite{michalik_analytic_2006}.
In the AG model the dynamics of a Gaussian electron pulse are computed using a mean-field self-similar approximation.
The result is a set of equations that can be solved very quickly using numerical methods on a modern computer.
When checked against N-body simulations, the authors find that the model is also surprisingly accurate, even for a large set of non-Gaussian pulses~\cite{michalik_analytic_2006,michalik_evolution_2009}.
A drawback is that the only force included in this model is that of the internal repulsive space-charge field.
In Chapter \ref{chap:extension}, I will extend this model, adding the influence of external forces, making it more generally applicable for UEM.

@@

For a classical set of $N$ discrete particles (labeled by the index $i$), distributed in 6D space (3 position, 3 momentum) the distribution function is 
\begin{equation}
  f_{\smallD}(\vec{r}, \vec{p}; t) = 
  \sum\limits^{N}_{i=1} \delta(\vec{r}_{i} - \vec{r}) \delta(\vec{p}_{i} - \vec{p}) \,\text{,}
\end{equation}
which automatically satisfies \ref{eq:normalization}.
If the particles have mass $m$ which interact with some potential $V(\vec{r}_{i} - \vec{r}_{j})$, their motion is governed by the equations
\begin{subequations}\label{eq:std_motion}
\begin{gather}
  \frac{d\vec{r}_{i}(t)}{dt} = \frac{\vec{p}_{i}(t)}{m}\\
  \frac{d\vec{p}_{i}(t)}{dt} = -\sum\limits_{j} \frac{d}{d\vec{r}_{i}} V(\vec{r}_{i} - \vec{r}_{j}) \label{eq:std_motion_v} \,\text{.}
\end{gather}
\end{subequations}
Now let us consider the time evolution of $f_{\smallD}$;
\begin{equation}
  \frac{\partial f_{\smallD}}{\partial t} = 
  \frac{d\vec{r}}{dt} \frac{\partial f_{\smallD}}{\partial \vec{r}} 
  + \frac{\partial f_{\smallD}}{\partial \vec{p}} \frac{d\vec{p}}{dt} \,\text{,}
\end{equation}
but with \ref{eq:std_motion} the final discrete evolution equation is
\begin{equation} \label{eq:evolution_discrete}
  \frac{d}{dt} f_{\smallD}(\vec{r}, \vec{p}; t) =
  -\frac{\vec{p}}{m} \frac{\partial}{\partial \vec{r}} f_{\smallD}(\vec{r}, \vec{p}; t)
  + \frac{\partial}{\partial \vec{p}} f_{\smallD} (\vec{r}, \vec{p}; t)
  \frac{\partial}{\partial \vec{r}} \iint d\vecprime{r} d\vecprime{p} V(\vec{r} - \vecprime{r}) f_{\smallD}(\vecprime{r},\vecprime{p};t) \,\text{,}
\end{equation}
where the extra negation comes from the change of perspective of \ref{eq:std_motion_v} from the frame of the individual particle $i$ to the observer frame.

This analysis can be extended to continuous distributions by taking the ensemble average over \ref{eq:evolution_discrete}.
In order to simplify the resulting equation to a useful result, one applies the usual mean-field approximation
\begin{equation}
  \left < f_{\smallD} (\vec{r}, \vec{p}; t) f_{\smallD} (\vecprime{r}, \vecprime{p}; t) \right > = \left < f_{\smallD} (\vec{r}, \vec{p}; t) \right > \left < f_{\smallD} (\vecprime{r}, \vecprime{p}; t) \right > \,\text{,}
\end{equation}
so that the formal evolution of a continuous distribution is given by
\begin{equation} \label{eq:formal_evolution}
  \frac{d}{dt} f(\vec{r}, \vec{p}; t) =
  -\frac{\vec{p}}{m} \frac{\partial}{\partial \vec{r}} f(\vec{r}, \vec{p}; t)
  + \frac{\partial}{\partial \vec{p}} f (\vec{r}, \vec{p}; t)
  \frac{\partial}{\partial \vec{r}} \iint d\vecprime{r} d\vecprime{p} V(\vec{r} - \vecprime{r}) f(\vecprime{r},\vecprime{p};t) \,\text{.}
\end{equation}
Although this equation looks very similar to \ref{eq:evolution_discrete}, notice that it contains approximations, mean-field that has been seen and in practice the self-similarity of the distribution function, that the discrete description did not.
